---
title: "HW1_[LASTNAME]" 
author: "[YOUR NAME]" 
date: "9/13/2022"
output: html_document #this tells R to output an additional HTML version of this file when knitting. 
---

# Set up and load packages

## Code Chunks and Regular text in . rmd

**Note**: You can write plain text in these parts of the script. You can write R code in the code chunks below, which begin with three back ticks and “{r}” and end with three more back ticks.

### Example: what is 1 + 1?
Verbal response to the HW question goes here. 1 + 1 = 1.
```{r}
# this is a code chunk
# here goes the code you ran to answer the question.
1 + 1
```

You can insert your own code chunks with a keyboard shortcut (command + option + “I” on Mac, ctrl + alt + “I” on PC).

To run code in the code chunks, you can either select a line of code and hit control + enter (on Windows PCs) or command + enter (on Macs), or click the small green triangle at the top-left corner of the chunk.

## Creating headings in .rmd

# 1st Level Heading
This is 1st Level Heading.

## 2nd Level Heading
This is 2nd Level Heading.

### 3rd Level Heading
This is 3rd Level Heading.


## Install and Load required packages

To load packages, run the code chunk below by clicking the small green triangle in the corner of the chunk.
```{r setup, message=FALSE}
# "r setup" above tells R that this code chunk is the setup chunk of this .rmd file, so it will run before any other code in the document. "message=FALSE" prevents messages/output generated by code from appearing in the knitted file. These are not necessary, nor should they appear, in a regular code chunk.

library(psych) # for describe and describeBy
library(tidyverse) # for read_csv
library(tinytex) #allows for markdown to write a pdf document
```



# Part I. Working with Data in R

## 1. Read the data in and check them out.

### a.
```{r}

```

### b.
```{r}

```

### c.
```{r}

```


## 2. Obtain the following basic information about your data:
Now try create your own code chunks.

### a.


### b.


### c.


### d.


### e.


## 3. Centering pretest:
Now try organizing sub-headings by the structure of HW1 on yourself.


## 4. Recode the condition variable, ...


## 5. Create a single global measure of overall learning by averaging the posttest measures:


## 6. Standardizing post_con score:


## 7. Create a table of descriptive stats in a Word document

**Write the code that allowed you to fill in the table.**


## 8. Make a scatter plot ...


## 9. Use the command we learned in lab to generate the mean posttest score for each condition.


# Part II: Reading Questions (Chapters 1 and 2)
Answer the following questions in a few sentences directly in your RMarkdown file (not within a code chunk).

## 10. When fitting a model to data, the data analyst faces two conflicting goals. What are those two goals and why are they conflicting?


## 11. What are two consequences of using sum of squared errors rather than the sum of errors as a summary measure of model error?


# Part III: Reflection
Take a minute to reflect on the materials covered in this homework. 

## 12. Which parts did you understand pretty well?

## 13. Which parts appear to you the most unclear or difficult? 

## 14. How long did it take you to complete this assignment? Write out the number of hours.


## 15. Knit your file using the Knit button near the top of your RStudio window. 

If knitting is successful, RStudio will also output an HTML file of your homework. Please submit your RMarkdown file, your HTML file, and your Word doc for question 7 in Canvas.

If you have trouble knitting your file, please contact Emma or Ivette. Congrats on finishing HW 1!
