--- 
output: html_document 
editor_options:  
  chunk_output_type: console
--- 
 

# Unit 15: The Generalized Linear Model

::: {.callout-important}
# Question

What are the assumptions of **general** linear models?
:::


:::{.fragment}
The general linear model makes the 5 assumptions below. When these assumptions are met, OLS regression coefficients are MVUE (Minimum Variance Unbiased Estimators) and BLUE (Best Linear Unbiased Estimators). 


1. **Exact X**: The IVs are assumed to be known exactly (i.e., without measurement error).
2. **Independence**: Residuals are independently distributed (prob. of obtaining a specific observation does not depend on other observations).
3. **Normality**: All residual distributions are normally distributed.
4. **Constant variance**: All residual distributions have a constant variance, $SEE^2$.
5. **Linearity**: All residual distributions (i.e., for each $Y'$) are assumed to have means equal to zero.
:::

----

::: {.callout-important}
# Question

What are the consequences of violating each of these assumptions? What options exist when these assumptions are violated?
:::


:::{.fragment}
1. **Exact X**: Biased parameters (to the degree that measurement error exists). Use reliable measures.
2. **Independence**: Inaccurate standard errors, degrees of freedom and significance tests. Use repeated measures or linear mixed effects models or ANCOVA.
3. **Normality**: Inefficient (with large N). Use power transformations, **generalized linear models**.
4. **Constant variance**: Inefficient and inaccurate standard errors. Use power transformations, SE corrections, weighted least squares, **generalized linear models**.
5. **Linearity**: Biased parameter estimates. Use power transformations, polynomial regression, **generalized linear models**.
:::

----

## Generalized Linear Models

```{r}
#| echo: false
#| message: false

library(tidyverse)
library(skimr)
theme_set(theme_classic()) 
```


```{r}
#| eval: false

glm(formula, family = familytype(link = linkfunction), data = data)
```

```{r}
#| echo: false


tibble(Family = c("binomial", "gaussian", "Gamma", "inverse.gaussian", 
                  "poisson", "quasi", "quasibinomial", "quasipoisson"),
       `Default Link Function` = c('(link = "logit")', '(link = "identity")', 
                                   '(link = "inverse")', '(link = "1/mu^2")', 
                                   '(link = "log")', 
                                   '(link = "indentity", variance = "constant")', 
                                   '(link = "logit")', '(link = "log")')) |> 
  kableExtra::kable() |> 
  kableExtra::kable_styling()
```

----

## An Example

Predicting admission to a grad program in engineering based on quantitative GRE, GPA, and Undergraduate Institution Rank.   

```{r}
#| echo: false

data <- read_csv(here::here("data_lecture/15_glm_admit.csv"),
                 show_col_types = FALSE)
```

```{r}
data |> 
  skim() |> 
  focus(n_missing, numeric.mean, numeric.sd, min = numeric.p0, max = numeric.p100) |> 
  yank("numeric")
```

----

```{r}
m_lm <- lm(admit ~ gpa, data = data)

broom::tidy(m_lm)
```


::: {.callout-important}
# Question

What is the effect of GPA on admission to grad school?
:::

----

::: {.callout-important}
# Question

What are the problems with using a general linear model to assess the effects of these predictors on admission outcomes?
:::

:::{.fragment}

1. Residuals will not be normal (not efficient).

2. Residual variance often will not be constant (not efficient, SEs are inaccurate).

3. Relationship will not be linear (parameter estimates biased).

4. Y is not constrained between 0 – 1 (model may make nonsensical predictions).
:::

----

```{r}
data |> 
  ggplot(aes(x = gpa, y = admit)) +
  geom_point() +
  scale_x_continuous(limits = c(0, 4))
```

----

```{r}
data |> 
  ggplot(aes(x = gpa, y = admit)) +
  geom_jitter(width = 0, height = .05) +
  scale_x_continuous(limits = c(0, 4))
```

----

```{r}
data |> 
  ggplot(aes(x = gpa, y = admit)) +
  geom_jitter(width = 0, height = .05) +
  scale_x_continuous(limits = c(0, 4)) +
  geom_smooth(formula = y ~ x, method = "lm", color = "red", se = FALSE)
```

----

```{r}
data |> 
  ggplot(aes(x = gpa, y = admit)) +
  geom_jitter(width = 0, height = .05) +
  scale_x_continuous(limits = c(0, 4)) +
  scale_y_continuous(limits = c(-2, 1.2)) +
  geom_smooth(formula = y ~ x, method = "lm", color = "red", se = FALSE)
```


----


```{r}
gvlma::gvlma(m_lm)
```





