---
editor_options: 
  chunk_output_type: console
---

# Simulate SEs for Logistic Regression {.unnumbered}


## Setup 
```{r}
#| messages: false
#| warnings: false

# packages
options(conflicts.policy = "depends.ok")
library(tidyverse)
library(furrr)

# source  
source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
source("https://github.com/jjcurtin/lab_support/blob/main/my_skim.R?raw=true")


# options
theme_set(theme_classic()) 
```

## Functions 

```{r}
make_data <- function(n, b0, b1){
  x <- runif(n, -10, 10) 
  z <- b0 + b1 * x
  p <- 1 / (1 + exp(-z))
  y <- rbinom(n, 1, p)
  tibble(y = y, x = x)  
}
```

```{r}
get_preds <- function(sim, n, b0, b1){
  d <- make_data(n, b0, b1)
  m <- glm(y ~ x, data = d, family = binomial) 
  preds <- tibble(x = c(-8, -3, 0, 3, 8))
  preds <- preds |> 
    bind_cols(predict(m, newdata = preds,
                      type = "link",
                      se.fit = TRUE)) |> 
    select(x, se_lo = se.fit, fit_lo = fit)
  preds |> 
    bind_cols(predict(m, newdata = preds, 
                      type = "response",
                      se.fit = TRUE),
           sim = sim) |>
    select(sim, x, fit_lo, fit_pr = fit, se_lo, se_pr = se.fit)
}
```

## Simulate sampling distribution

Settings
```{r}
# set n high to have robust models (and to handle uncertainty about z vs. t)
n <- 1000 
b0 <- 0
b1 <- 0.5 
n_sims <- 50000
```

simulate predictions to create sampling distribution
```{r}
cl <- parallel::makePSOCKcluster(parallel::detectCores(logical = FALSE))
doParallel::registerDoParallel(cl)
plan(multisession, workers = parallel::detectCores(logical = FALSE))

preds <- 1:n_sims |> 
  future_map(\(sim) get_preds(sim, n, b0, b1),
             .options = furrr_options(seed = 2468)) |>
  list_rbind()

plan(sequential)
```

## Add post simulation calcs to preds

Transform log-odds to pr and get CIs for log-odds, pr from log-odds, and native pr
```{r}
preds <- preds |> 
  mutate(upr_lo = fit_lo + 1.96 * se_lo,
         lwr_lo = fit_lo - 1.96 * se_lo,
         fit_lo_pr = plogis(fit_lo),
         upr_lo_pr = plogis(upr_lo),
         lwr_lo_pr = plogis(lwr_lo),
         upr_pr = fit_pr + 1.96 * se_pr,
         lwr_pr = fit_pr - 1.96 * se_pr) 
```

Calc true values and join with preds 
```{r}
true_values <- tibble(x = c(-8, -3, 0, 3, 8)) |> 
  mutate(true_lo = b0 + b1 * x,
         true_pr = plogis(true_lo))

preds <- preds |> 
  left_join(true_values, by = "x") |> 
  relocate(sim, x, true_lo, true_pr, 
           fit_lo, lwr_lo, upr_lo, 
           fit_lo_pr, lwr_lo_pr, upr_lo_pr,
           fit_pr, lwr_pr, upr_pr)
```

Calculate true SE (sd of sampling distribution for fits across simulations)
```{r}
se_true <- preds |> 
  group_by(x) |>
  summarise(se_lo_true = sd(fit_lo), 
            se_pr_true = sd(fit_pr),
            .groups = "drop")
```

Review preds
```{r}
preds |> head(10)
```

## Visual sampling distributions for predictions

For log-odds
```{r}
preds |> 
  ggplot(aes(x = fit_lo, fill = fct(as.character(x)))) +
  geom_density(alpha = 0.5) +
  labs(title = "Sampling Distribution of Log-Odds",
       x = "Log-Odds",
       y = "Density") + 
  scale_fill_discrete(name = "X")
```

For pr from log-odds
```{r}
preds |> 
  ggplot(aes(x = fit_lo_pr, fill = fct(as.character(x)))) +
  geom_density(alpha = 0.5) +
  labs(title = "Sampling Distribution of Pr from Log-Odds",
       x = "Pr",
       y = "Density") + 
  scale_fill_discrete(name = "X")
```

For pr
```{r}
preds |> 
  ggplot(aes(x = fit_pr, fill = fct(as.character(x)))) +
  geom_density(alpha = 0.5) +
  labs(title = "Sampling Distribution of Pr",
       x = "Pr",
       y = "Density") + 
  scale_fill_discrete(name = "X")
```

Look more closely at edge of X.  Its mostly but not completely symetric.   This is why a symetric confidence interval will NOT work optimally for pr. 
```{r}
preds |> 
  filter(x == -8) |> 
  ggplot(aes(x = fit_pr)) +
  geom_density(alpha = 0.5) +
  labs(title = "Sampling Distribution of Pr for X = -8",
       x = "Pr",
       y = "Density")
```


## Review min and max values for probability based CIs

CIs for probability are NOT equivalent when formed from log-odds vs. pr.  This is because the transformation from log-odds to pr is non-linear (after adding/subtracting SE).

In this example, the CI boundaries for probability are never < 0 or greater than 1 even when calculated directly as probability + 1.96 SE.  But that may  not be always true.  Also clear that +- 1.96 SE is not appropriate for a probability CI given its not symetric and therefore clearly not always normal.
```{r}
preds |> 
  group_by(x) |> 
  summarise(min_lwr_lo_pr = min(lwr_lo_pr),
            min_lwr_pr = min(lwr_pr),
            max_upr_lo_pr = max(upr_lo_pr),
            max_upr_pr = max(upr_pr),
            .groups = "drop")
```

## Check model SEs (from predict) for lo and pr

Now fit a single model and use `predict()` to se.fit for lo and pr 
```{r}
d <- make_data(n, b0, b1)
m <- glm(y ~ x, data = d, family = binomial) 
sample_preds <- tibble(x = c(-8, -3, 0, 3, 8))
sample_preds <- sample_preds |> 
  bind_cols(predict(m, newdata = sample_preds, 
                        type = "link",
                        se.fit = TRUE)) |> 
  select(x, se_lo_mod = se.fit)
sample_preds <- sample_preds |> 
  bind_cols(predict(m, newdata = sample_preds, 
                        type = "response",
                        se.fit = TRUE)) |> 
  select(x, se_lo_mod, se_pr_mod = se.fit)
```

Now compare model SEs to true SEs

```{r}
se <- se_true |> 
  full_join(sample_preds, by = "x") |> 
  mutate(diff_lo = se_lo_true - se_lo_mod,
         diff_pr = se_pr_true - se_pr_mod) |> 
  relocate(x, 
           se_lo_true, se_lo_mod, diff_lo, 
           se_pr_true, se_pr_mod, diff_pr)
se
```

## Check CIs

For LO
GOOD!
```{r}
preds |> 
  mutate(value_in_ci = if_else(lwr_lo <= true_lo & 
                               upr_lo >= true_lo, 
                               1, 0)) |> 
 pull(value_in_ci) |> 
 mean()
```


For PR from LO 
GOOD!
```{r}
preds |> 
  mutate(value_in_ci = if_else(lwr_lo_pr <= true_pr & 
                               upr_lo_pr >= true_pr, 
                               1, 0)) |> 
 pull(value_in_ci) |> 
 mean()
```

For PR from LO 
Not perfect.  This is why we should use CIs for probability formed from log odds predictions!
```{r}
preds |> 
  mutate(value_in_ci = if_else(lwr_pr <= true_pr & 
                               upr_pr >= true_pr, 
                               1, 0)) |> 
 pull(value_in_ci) |> 
 mean()
```


As expected the CI coverage is worse for the edge of the X range
```{r}
preds |> 
  filter(x == -8) |>
  mutate(value_in_ci = if_else(lwr_pr <= true_pr & 
                               upr_pr >= true_pr, 
                               1, 0)) |> 
 pull(value_in_ci) |> 
 mean()
```