---
title: "hw_06_shell" 
author: "TAs" 
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---


## Part A: Conceptual Questions

### 1. A researcher presents the following equation representing the relationship between the number of books a child reads and whether or not that child has a diagnosis of dyslexia, where dyslexia is coded 1 (does not have dyslexia) and -1 (has dyslexia): Books = 1.52 + 0.5 * dyslexia. If the researcher recoded dyslexia as -.5 (does not have dyslexia) and .5 (has dyslexia), what would be the new regression equation representing the predicted number of books a child reads?

> Your answer here.

### 2. What does it mean for an estimate to be "unbiased"? What is our estimate for the population mean? Is this estimate biased or unbiased? What is our estimate for the population SD?

> Your answer here.


## Part B: Data Analysis

Researchers at UW-Madison conduct a study about laughing. They have a series of hypotheses about the social function of laughter, but for this homework you will focus on only one of these hypotheses: **that alcohol consumption has a positive relationship with amount of laughing**.     
 
After a home football game, researchers recruited both Badger fans and the rival team’s fans in pairs as they were exiting Camp Randall Stadium, and brought them to a lab in Brogden. Half of the subjects were paired with a fan of their own team (whom they had never met before), and half were paired with a fan of the rival team. Since gender differences exist in average laughter frequency, all participants were paired with same-gender partners (Smoski & Bachorowski, 2003). Participants were not given an opportunity to talk to each other before the study began. The subject pairs were seated together in front of a TV and watched 7 video clips from America’s Funniest Home Videos. The researchers filmed the session so they could code the total duration of participants’ laughter that occurred during each video clip. After the video viewing stage, the participants separately completed scales measuring how much they liked and how similar they felt to their partner and reported how many drinks they had had that day (since the researchers expected more intoxicated participants to laugh more). The researchers didn’t know how to handle the non-independence of the subject pairs, so they just computed average scores for each pair on all the variables. The codebook for the data (`hw_06_data.csv`) is given below.   

Variables:   

- `team_id`:	Unique identifier for each pair of participants.
- `team`:	Whether participant pairs were fans of the same team (`same`) or rival teams (`different`).	   
- `laughter_1` - `laughter_7`: The total duration (in seconds) of laughter observed during each of the 7 video clips	(possible range = 0 - 23.75 seconds).
- `liking`:	Mean/composite score of a scale measuring how much the participant pairs liked each other at the end of the study (0 - 7 scale, 0 = strongly dislike, 7 = strongly like).
- `similarity`:	Mean/composite score of a scale measuring how similar the participants found their partners to be (averaged for each pair, 0-7 scale, 0 = very dissimilar, 7 = very similar).
- `drinks`:	Number of drinks consumed that day (averaged for each pair).



### 1. Set up your script, then read in the `hw_06_data.csv` data and inspect them. Based on descriptive statistics and univariate plots, are there any hints that some cases might have been miscoded and should be excluded?

```{r}
#| message: false

# set conflict policy, import packages, and set up graph theme


# set path for data


# source var_score function

```

Read in Data
```{r}
d <- 
```

Inspect Data
```{r}

```


There was definitely some sort of error with the `drinks` variable, which has a max of 7,000. Yikes! Let’s see the distribution of `drinks`.

```{r}

```

It looks like there’s just one person who is an extreme outlier. We’ll want to exclude this case. I’m a bit suspicious that something is wrong with the `laughter_6` variable, because the max is so low. Let’s look at its distribution too.

```{r}

```

The modal laugh time was 0 seconds here, and those who did were mostly laughing for less than 1 second. Maybe this was just a really non-funny clip, worthy of only a single brief chuckle for some people.


### 2. Observations in this dataframe should have unique identifiers. Create a variable called `pair_id` (an id for each observation in the dataset) that ranges from `p1001` to `p1209`. Reorder your variables so this id is in the first column. Google it if you need help with any of this data wrangling. `str_c()`, and `relocate()` are two functions that might help you get the job done.

```{r}
d <- 
```


### 3. Exclude any cases that have clearly been miscoded. While in your own data you might look further into this miscoding to see if you can fix it, for the purposes of this assignment, simply remove the entire pair when you find evidence of clear miscoding. Do a quick check for any additional miscoded values.

Descriptives look much more as expected when removing the outlier from the `drinks` variable. 
```{r}
# remove the outlier
d <- 

# quick check on `drinks`

```

Let’s check the other variables to see if there are any other miscoded values.   
Ranges are within expected values for the other variables.   
```{r}
# check team


# check liking & similarity


# check variables starting with `laughter`

```


### 4. Focus on the variables measuring laughter during each video clip (i.e., `laughter_1` - `laughter_7`). After doing some digging, you found out that your researchers did not use the standardized version of the laughter procedure and accidentally showed pairs a (very sad) clip from the opening of the Pixar movie “Up,” which is why `laughter_6` looked different than the others. Create a composite measure of laughter (`laughter_m`), excluding this anomalous clip.

```{r}
d <- d |> 
  mutate(laughter_m = )
```

### 5. Explore the dataframe by obtaining the bivariate correlations and scatterplot matrix on some of the variables of interest: `laughter_m`, `liking`, `similarity`, and `drinks`. Round the results to two digits. Note any strong relationships between these variables.

```{r}
#| message: false

d |>
  


```

Is there any strong relationship? Plese specify.

> Your answer here.

### 6. Fit a model in which you predict the average laughter from drinking. In a sentence, describe your findings (report F, p values, and partial eta squared). Create three functions that calculate SSE, PRE, and the F-statistic.

Let's first create a few functions. Maybe you already had these saved and you sourced them in your own function script? If not, create three functions that calculate sse, pre, and f_stat.
```{r}
sse <- 

pre <- 

f_stat <- 
```

Find F, p values, and partial eta squared
```{r}
m <- 


```

Report your results

> Your answer here.

### 7. What is the interpretation of the intercept in this model? Is this value meaningful?

> Your answer here.

### 8. Make a publication quality plot of the relationship between drinks and laughter. Below the plot, create a caption that describes what the error bands in your plot represent.

```{r}
data_new <- 

preds <- 

ggplot()
```


### 9. Fit a model in which you predict the average laughter from drinking when controlling for the effect of similarity. In a sentence, describe your findings (report F, p values, and partial eta squared).
```{r}
m_2 <- 


```

Describe your findings.

> Your answer here.

### 10. Provide an interpretation for each of the parameter estimates for the model you fit in question 9. Your answer must include a $b$ value and a description that can be understood by someone who is not a statistician.

> Your answer here.


## Part C: Reflection

### How many hours did this assignment take you?

> Your answer here.

### What is one conceptual topic you would like us to spend extra time reviewing in preparation for your in-class concepts exam?

> Your answer here.

### Take a moment to reflect on having completed your first application exam (congratulations!). What went well? What do you feel could have gone better, and how can we best support you to prepare for the following application exams?

> Your answer here.

### Free space to put any other comments about lab and/or lecture (but not required!).

> Your answer here.

