---
title: "hw_12_key" 
author: "TAs" 
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---


## Part A: Conceptual Questions

### 1.	In the textbook, what do Judd and colleagues recommend in terms of reporting *simple* and *main* effects? What is their justification? Does Jaccard (2003) agree or disagree with this perspective? 



### 2.	According to MacCallum and colleagues (2002), when should you dichotomize a continuous variable? Why?



### 3.	A researcher wants to determine if living in an urban versus rural area predicts anxiety. The researcher goes door to door in two towns (one urban and one rural) and asks all people living in each dwelling (home, apartment, etc.) questions to assess their anxiety. In multiple people live in the same dwelling, the researcher asks each of them to answer the questions. In two or three sentences, explain which model assumption is most directly violated by this research design.



## Part B: Data Analysis

Some researchers plan to examine whether alcohol consumption affects dart playing skills. Their study aims to explore the well-known placebo effect: if people believe they are consuming alcohol, or another drug, they may feel and behave as if they are actually under the influence of that substance when in reality they are not. Thus, the purpose of the study is to clarify whether it is the effect of alcohol or the expected effect of alcohol that leads to differences in performance.

Sixty dart players are invited to an open-bar event where dartboards are made available. Participants were randomly assigned to either receive real alcohol or a convincing placebo. Non-alcoholic placebos were piloted beforehand and shown to be indistinguishable from alcoholic beverages. Players reported their prior experience with dart games and self-assessed coordination (rated right before the game) to control for differences in dart playing ability and to monitor for interactions with the experimental manipulation. Performance scores during the study ranged from 0-150. Blood alcohol concentration (BAC) was also determined. BAC was multiplied by 100 in this dataset.

Researchers want to test their confirmatory hypothesis that the effect of BAC on game scores would be stronger for less experienced players. In addition, they also want to run exploratory analysis to test whether BAC and self-rated coordination interact to predict game scores.


**Codebook:**   

- `sub_id`: unique subject ID. (1 - 60) 
- `condition`: alcohol condition. (placebo; alcohol)    
- `experience`: prior experience with dart games. (novice; experienced)   
- `bac`: blood alcohol concentration * 100. (0 - 10)   
- `coordination`: self-assessed coordination. (0/poor - 10/good)   
- `score`: performance scores during the study. (63 - 150)   

```{r}
#| message: false

```

### 1. Data preparation

#### a. Read in the data file (`hw_12_data.csv`). 

```{r}

```


#### b. Explore univariate descriptive statistics for all variables. Also calculate bivariate correlations among continuous variables.

```{r}

```

#### c. To test all three hypotheses, what variables should be centered to aid the interpretation of your models? Why? 


#### d. Create centered versions of the variables you named in question 1c.
```{r}

```


### 2.	Group the data by all possible combinations of `condition` and `experience`. Then, run code to obtain the means and standard deviations for `score`, `coordination`, and `bac` in each of these groups.
```{r}

```


### 3. Does the effect of BAC on game scores depend on experience?

#### a.	Fit a model that addresses this question.
```{r}

```


#### b.	Report and interpret each of the parameter estimates. Report F-value, and p-value for non-intercept estimates (one sentence each is sufficient).





#### c.	Create a publication quality graph depicting the first model that you fit.

```{r}

```

#### d.	Fill in the blanks for this regression equation, recording your answers as A-D in your QMD document:
score = ___A___ + __B___ * bac_c + __C___ * experience_c + ___D____ * bac_c * experience_c


#### e.	What is the magnitude of the effect of BAC for novice participants? Find the answer using the equation above, without recoding your variables in R.



#### f.	What is the magnitude of the effect of experience groups for an individual with a BAC of 0? Find the answer using the equation above, without recoding your variables in R.



#### g.	Test the simple effect of BAC for individuals in the novice group by fitting a model with experience groups dummy coded. Interpret the relevant coefficient and report an F-value and p-value.

```{r}

```



#### h.	Test the simple effect of experience for someone with a BAC one standard deviation above the mean of that variable by fitting a model with recentered BAC. Interpret the relevant coefficient and report appropriate statistics.

*Note:* to recode so 1 SD **above** the mean is equal to 0, you **subtract** 1 SD from the mean-centered version. If you add 1 SD to the mean-centered version, that would make 1 SD below the mean = 0 (NOT what you want!)

```{r}

```



#### i. Run code to conduct case analysis on the interaction model. 
```{r}


```


#### j. Run code to examine if there are any violations to the assumption of normality and the assumption of constant variance. 

**Note: Skip the assumption of linearity for now, since the component+residual plots do not accommodate interaction terms.**

```{r}

```

#### k. Summarize the findings from your case analysis and inspection of model assumptions. Are there any issues? What can potentially be done to address these issues?



```{r}

```


### 4.	Does the effect of BAC on game scores depend on coordination?

#### a.	Fit a model that addresses this question.

```{r}

```

#### b.	Report and interpret each of the parameter estimates, reporting relevant statistics (one sentence each is sufficient).


#### c.	Test whether there is a significant simple effect of coordination for someone who has a BAC at the legal limit (coded as 8 in this data file) by fitting a model with recentered BAC. Interpret the relevant coefficient and report appropriate statistics.

```{r}

```


### 5.	What would happen to the SE of the parameter estimate of the interaction term in the previous model if you had used the uncentered versions of blood alcohol concentration and coordination? Would it increase, decrease, stay the same, or is it impossible to say what will happen to it?


### 6.	How long did this assignment take you (in hours)?



