---
title: "hw_11_shell" 
author: "TAs" 
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---


## Part A: Conceptual Questions


### 1. Of the assumptions of linear regression that refer to residuals, which two are the most problematic to violate in large samples? Why are these particularly concerning? What are solutions to violating these assumptions?




### 2. Explain the logic of a Q-Q plot and what normally distributed data would look like (about 3 sentences).




## Part B: Data Analysis

A public health organization wants to test the importance of staying active and healthy in the winter by examining links between cross-country skiing, vegetable consumption, and health, as measured by an independent doctor at the end of winter. Participants were 53 residents of the Madison, WI area who were recruited at the beginning of the winter and asked to record the number of times they went cross-country skiing over the course of the winter and their average daily vegetable consumption (in servings). At the end of the winter (i.e., the last week of March), a doctor conducted a health assessment of each participant and reported their health on a 0-10 scale, with 0 representing very poor health and 10 representing excellent health.    

The researchers in the public health organization predict that cross-country skiing and vegetable consumption both contribute to overall health. They also expect to see little to no relationship between vegetable consumption and skiing behavior.   

The following variables can be found in the `hw_11_data.csv` dataset.    

- `sub_id`: Unique subject identification number (range 1 - 53).
- `skiing`:	Number of times participant went cross-country skiing over the course of the winter (range 0 - 93).
- `veggies`: Number of servings of vegetables participant consumes on an average day (range 0 - 16).
- `fitness`: Level of health, as assessed by doctor, at the end of March (range 0-10).


### Setup
```{r}
#| message: false





```

### 1. Read in and inspect the data. Also source the functions you will use for case analysis.

```{r}


```


### 2. Fit a model testing the researchers’ hypothesis. Specifically, run a model testing whether skiing and vegetable consumption both uniquely explain variance in people’s health scores at the end of March (when included in the same model). Report your results (F stats, p-values, and model $R^2$) in a sentence.

```{r}


```


### 3. Examine whether there are violations to model assumptions for the model you fit above.

#### a. Is the assumption of normality violated? How do you know?

```{r}


```

#### b. Is the assumption of constant variance violated? How do you know?

```{r}


```

#### c. Is the assumption of linearity violated? How do yoy know?

```{r}


```


### 4. You suspect there may be outliers present in the dataset. 

#### a. Identify which participants have high leverage. Note the variables on which each of these participants is extreme (i.e., why it has a large hat value). 

```{r}


```


#### b. Identify participants that are regression outliers. Why are these participants regression outliers? 

```{r}


```


#### c. Examine the Cook’s D influences of the participants. Are the participants who have extreme Cook’s D scores surprising to you or not?

```{r}


```


### 5. Remove the 4 participants who have the highest Cook's D influence from the data and run your focal model again. What has changed? Why?

```{r}


```


### 6. Examine whether there are violations to model assumptions for the model you just fit with outliers (i.e., high influence points) removed. 

#### a. Is the assumption of normality violated? How does this model compare to the model above before removing the outliers, in terms of residual normality?

```{r}


```

#### b. Is the assumption of constant variance violated? How does this model compare to the model above before removing the outliers, in terms of constant variance?

```{r}


```

#### c. Is the assumption of linearity violated? How does this model compare to the model above before removing the outliers, in terms of linearity?

```{r}


```


### 7. In the exercise above, we examined model assumptions first, conducted case analysis, and then examined model assumptions again with outliers removed. What is your take away from this exercise? How does case analysis influence your inspection of model assumptions? Write 2-3 sentences to reflect.




### 8.	How many hours did this homework take you?

