---
title: "Lab 12 TA Script" 
author: "TAs" 
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

## Objectives

Welcome to lab 12!

Today we will cover:
1. Conceptual exercise: understanding effects from a regression equation through manual calculations.
2. Continuous-by-continuous interactions
3. 2-by-2 ANOVA

## Conceptual Exercise

Suppose you sampled 400 participants and asked them to report whether or not they read for pleasure (a score of -0.5 indicates they do not read for pleasure and a score of 0.5 indicates they do read for pleasure), the number of hours spent outside per week (ranging from 0 to 20), and happiness (measured on a 50-point scale with greater numbers indicating greater happiness). There are the same number of people who read for pleasure as there are people who do not read for pleasure. On average, participants in your sample spend 10 hours outside per week. You obtain the following results:    

$happiness = 10 + 3 * read_c + 2 * outside_c + 0.5 * read_c * outside_c$   


### 1. What is the magnitude of the simple effect of time spent outside on happiness for participants who read for pleasure?







### 2. What is the magnitude of the simple effect of time spent outside on happiness for participants who do not read for pleasure?   







### 3. What is the effect of reading on happiness for participants with an average number of hours spent outside? Is this a simple or main effect?   







### 4. Describe in words how the effect of reading on happiness varies by the number of hours spent outside.    




### 5. How many hours does a person spend outside if they have a score of -3 on the outside_c predictor?   




### 6. What is the predicted happiness for a person who spends 14 hours outside per week and does not read for pleasure?   





### 7. Change the coding of reading to 0 = does not read and 1 = reads for pleasure and fill in the resulting equation:    

    happiness = ___ + ___ * read + ___ * outside_c + ___ * read * outside_c    




### 8. How (if at all) will the standard errors for the parameter estimates for the interaction terms of the two models (one with read_c and one with read) differ?  




### 9. A lab member suggests to you that you can test whether you have a significant interaction by examining the simple effects of spending time outside at each level of reading for pleasure. That lab member tells you that if one is significant and the other is not, you will have a significant interaction. Do you agree with this lab member? Why or why not?  




### 10. Write the augmented and compact model for the model comparison which tests whether the read_c * outside_c interaction effect is significant.    

    Compact: 
    Augmented: 


**End of exercise!**   

`CHECKPOINT`

## Setup 

```{r}
#| message: false

options(conflicts.policy = "depends.ok") 
library(tidyverse)
library(broom)
library(skimr)

theme_set(theme_classic()) 

path_data <- "lab/data"
```

## Continuous x continuous interaction

**We will use the data from last week again. Read in `lab_11_data.csv`**

Remember, the data was from a hypothetical study where the researcher examined whether preschoolers benefit from collaborative learning (vs. passive learning) when they are learning about the external world. We also tested whether older preschoolers learn better in collaboration, such that the benefit would be considerably smaller for younger preschoolers. We had three variables:

- `age` (continuous; months since birth)
- `condition` (dichotomous; Observe [0] or Collaborate [1])
- `learning` (sponge uses in percentage indicative of preschoolers' learning)

In addition, parents of the preschooler also indicated the extent to which the preschooler engage in social behaviors that are intended to benefit others (e.g., sharing).

- `prosociality` (7-point Likert scale; a higher score indicates being more prosocial)   

###  11. Within the collaborate condition, it is possible that preschoolers benefit from developmental maturity (i.e., being older) only if they like interacting with and helping other kids (i.e., are prosocial). Thus, the researcher now wants to test the below exploratory hypothesis: *within the Collaborate condition*, the effect of age is moderated by preschooler prosociality. Load in the data.

```{r}
d <- read_csv(here::here(path_data, "lab_11_data.csv"),
              show_col_types = FALSE) |> 
  glimpse()
```


### 12. Explore the univariate statistics and distributions related to `age` and `prosociality`.

```{r}



```

### 13. Subset the data and get observations in the Collaborate condition only.

```{r}

d_collab <- d |> 
  
```

### 14. Prepare your data for analysis.

```{r}
d_collab <- d_collab |> mutate(

  
  )
```

### 15. Fit the appropriate model to test the hypothesis.

```{r}
m_1 <- lm()
```


### 16. Interpret the coefficients.

```{r}
tidy(m_1)
```

1. Intercept: b0 = 0.417  
The preschoolers who are of average age (M_age = `r mean(d_collab$age)` in months) and and average prosociality (M_prosociality = `r mean(d_collab$prosociality)`) are predicted to use the sponge to help clean up the water on 41.7% of the trials.

2. age_c: b = 0.025     
Preschoolers with average prosociality are predicted to use the sponge to help clean up the water on 2.5% more trials if they are 1 month older, controlling for the effects of prosociality and its interaction with age.

3. prosociality_c: b = 0.022   
Preschoolers with average age are predicted to use the sponge to help clean up the water on 2.2% more trials if they scored 1 point higher on the prosociality scale, controlling for the effects of age and its interaction with prosociality.

4. interaction: b = 0.013   
The effect of age is stronger in magnitude for preschoolers with higher prosociality, such that the increased likelihood of using the sponge to help is associated with being 1 month older is 1.3 units larger if the preschooler scored 1 point higher on the prosociality scale. 

OR

The effect of prosociality is stronger in magnitude for older preschoolers, such that the increased likelihood of using the sponge to help is associated with scoring 1 point higher on the prosociality scale is 1.3 units larger if the preschooler is 1 month older. 

### 17. Create a publication-quality graph for this model.

```{r}

















```

`CHECKPOINT`

## 2 x 2 ANOVA   

**Clean up your R environment.**

```{r}
rm(list = ls())
```

A researcher is testing whether watching a video about the benefits of a new tax policy moderates the relationship between political affiliation and support for that policy. The researcher recruits participants and randomly assigns them to either watch or not watch a short video. The researcher then measures whether they support the policy and determines the participant's political affiliation with a continuous and dichotomous measure, respectively.   

The codebook for this dataset:   
- `cond`: Assigned condition; 0 = no video, 1 = video    
- `support`: Support for the new tax policy;  0 (do not support at all) to 10 (strongly support)   
- `political_dichot`: Political affiliation measured dichotomously; 0 (Republican), 1 (Democrat)   


### 18. Run the code below to simulate this dataset.
```{r}
set.seed(123)

d <- tibble(cond = sample(0:1, 600, replace = TRUE),
            political_cont = sample(1:7, 600, replace = TRUE)) |>
  mutate(political_dichot = if_else(political_cont %in% c(1, 2, 3), 0,
                                    if_else(political_cont %in% c(5, 6, 7), 1, 
                                            sample(0:1, 600, replace = TRUE))),
         interact = political_cont * cond,
         support = round(pmax(0, pmin(10, rnorm(600, mean = 5, sd = 2) + 0.2 * interact)))) |> 
  select(-interact, -political_cont) |> 
  glimpse()
```

### 19. Examine the univariate and bivariate summary statistics for the variables in the dataframe, `d`.
```{r}





```


### 20. Create two tables reporting the number of people at each level of the predictor variables (`cond` and `political_dichot`).
```{r}


```

### 21. Create a table reporting the average support score for each level of the predictor variables.
```{r}





```

### 22. Create a contingency table reporting the average support score for all combinations of condition and political affiliation (measured dichotomously).    

```{r}







```

Based on the table you created, are the means consistent with a model in which the interaction between `cond` and `political_dichot` is significant? How can you tell?   

The simple effect of political affiliation differs at each level of condition as do the simple effects of condition at each level of political affiliation. An interaction means that the magnitude of the simple effects differs. Thus, these data are so far consistent with the proposed interaction. However, we will not know for sure until we test the model.   

**If you were to fit a linear model such that:**
$support = b0 + b1*cond + b2*political + b3*cond*political$
**...what values will b0, b1, b2, and b3 be, by just looking at the table you created above?**

b0 = 
b1 = 
b2 = 
b3 = 


### 23. Fit a linear model to test whether condition and political affiliation (measured dichotomously) interact to predict support for the tax policy.
```{r}
d <- d |> 

m_1 <- lm()
tidy(m_1)

```


### 24. Create a publication quality graph bar graph depicting the cell means with error bars representing the standard error of the point estimates from the model you just tested.  
```{r}















```

We have covered how to use different methods to fill/color your figures, including using color names (for a select amount of colors) and using premade palettes. Here, we use `scale_fill_manual()` in order to add in our own HEX codes -- the color customization possibilities are endless!

### 25. Provide a 1-2 sentence description of the model you tested and comment on whether the interaction was significant.




### 26. Re-test this same model, but instead of using the centered predictors, use the uncentered predictors. What changed between the two models?
```{r}
m_2 <- lm()
tidy(m_2)
tidy(m_1)
```


