---
title: "Homework 3" 
author: "[YOUR NAME HERE]" 
date: "`r lubridate::today()`"
format: 
  html:
    toc: true 
    toc_depth: 4
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

Due <!--Friday, September 27, 2023, 9:00pm--> on Canvas.   

-	Select the Assignments menu and find HW Week 3
-	Submit both an HTML file and a .qmd file
-	Make sure you can access this folder ahead of the homework due date


## Part A. Reading Questions


### 1.	Define Type I and Type II error. For each type, give an example of how a researcher could reduce the risk of committing the error.




### 2.	In practical terms, what does the numerator of the F statistic equation represent? The denominator?



### 3.	Statistical terms: What special name is given to the mean-squared-error from the mean-only model?




## Part B. Data Analysis


For this homework, we will use the `stray_animals.csv` file. This data set contains (fictitious) statistics about the number of stray animals in each of the 50 US states. The data set also specifies whether the state participated in a nation-wide TNR (trap-neuter-return) program that was launched in an attempt to reduce the number of stray cats and the number of animal shelters in each state.   

**Codebook:**    

- `state`: US state
- `tnr`: whether the program adopted a trap-neuter-return program (0/no, 1/yes)   
- `dogs`: Number of stray dogs in thousands
- `cats`: Number of stray cats in thousands
- `other`: Number of stray other animals in thousands
- `shelters`: Number of animal shelters per 10,000 people


### 1. Set up your global options for your script. This includes setting the conflict policy, loading any libraries you need, selecting a theme to use for `ggplot`, and specifying the path to where you store your data files (i.e., `path_data`).

```{r}
#| message: false


```


### 2.	Read in and glimpse the data. In what variable classes are the `tnr`, `dogs`, `cats`, `other`, and `shelters` columns read into R? What are the appropriate variable classes for these variables, according to the data codebook? Do they match?

***Your Answer Here***

```{r}



```

### 3. Use one function/command in R to pull up the help page for `factor()`. After you have read the help page, recode `tnr` to be a factor variable with clear labels (yes/no). Feel free to consult other sources, such as ChatGPT.

```{r}



```


### 4. Obtain summary statistics for all numeric variables using `summarise()`.

```{r}



```


### 5. Obtain summary statistics by whether or not a trap-neuter-return program is adopted, using `group_by` and a function from `skimr`.

```{r}



```


### 6.	Go through the following steps to create two additional variables:

#### a.	Compute the total number of stray animals (i.e., dogs, cats, and other animals combined) as `stray_total` in each state.

```{r}



```

#### b. Create a new variable `stray_z`, that is the standardized score of `stray_total`.

```{r}



```


#### c.	Check to make sure `stray_z` is standardized. 

```{r}



```


### 7.	Use ggplot to create two histograms of the total number of stray animals: one for the states that took part in the TNR program, and the other for the states that did not. Label your histograms properly. 

There are a few different ways to do this. Try multiple ways if you rock (hint: look into  using `filter()`, `subset()`, or `facet_wrap()`)!

```{r}



```


### 8.	Refer to the ggplot2 cheat-sheet [here](https://rstudio.github.io/cheatsheets/html/data-visualization.html) and other sources like ChatGPT. Use `ggplot`to create a *scatterplot* that shows the relationship between total number of stray animals and number of shelters. The plot should at least have a white background, data points, and meaningful x- and y-axis labels. 

```{r}



```


### 9.	What is the formula for each of the two models you should compare if you want to know whether the average number of stray animals is different from zero?   

***Your Answer Here***


### 10.	Compute the SSE for each of the two models in the previous question. Do so by using a step-by-step brute force method, like we did in lab. Remember to label clearly (e.g., `sse_null`, `sse_mean`).  

Compact - NULL model
```{r}



```

Augmented - Mean-only model
```{r}



```


### 11.	Compare the two models by calculating the appropriate F statistic and p value. 

```{r}



```

### 12. Now, fit the null model and the mean-only model using `lm()`. Then, use a function we learned to compare the two models.

```{r}



```


### 13. Use a function from the `broom` package to obtain the parameter estimate, its standard error, and the corresponding t-value and p-value from the mean-only model.

```{r}


```


### 14. Answer the following questions:

#### a. Do you get the same conclusion and results from step-by-step brute-force model comparison, `lm()` model comparison, and the method using the augmented model only and the `broom` package? 

***Your Answer Here***

#### b.	Is estimating the mean worth the increased complexity of the model? Why? Interpret the result in plain English.     

***Your Answer Here***

#### c.	Why is this model comparison not super informative (i.e., why are you not surprised 0 is a worse predictor for the number of stray cats per year?).    

***Your Answer Here***

### 15. Think back to the standardized score `stray_z` which we created from `stray_total`. Would it make sense to estimate b_0 and test whether `stray_z`has a mean that is significantly different from zero? Why or why not?

***Your Answer Here***

### 16.	Sometimes learning stats or programming is difficult and you just need a little positive feedback. Figure out which package contains the `praise()` function, and complete all the steps you need in order to write a line of code using `praise()` to tell you something nice. 


```{r}



```


### 17. Reflection

#### a. How many hours did you spend on this homework?

***Your Answer Here***


#### b. What is one aspect of Friday labs that has been effective so far?

***Your Answer Here***


#### c. What is one way Friday labs can be improved to facilitate your learning?

***Your Answer Here***


#### d. What is one concept covered in lecture that you feel the most confusing/elusive?

***Your Answer Here***


